SAT Competition 2020\tnoteref{title}

### 2021-02-02 paper/main.tex
sects 1-2, added some notes in red...

The same rules for participation as in Main track applied to the Planning track, in which we evaluated the solvers on 200 benchmarks 
which all came from the same application domain -- automated planning. 
Solver submitted to the Main track automatically participated in the Planning track.
It is envisioned that also in the future there will be an ``application promotion'' track,
each time highlighting a different area where the SAT solving technology helps to advance the state of the art.

Finally, the Main Track also had the so-called ``Glucose hack'' sub-track. 
Since in the past several advances in SAT solving required only a small modification of an established solver
to achieve a considerable contribution, this sub-track encouraged participation 
of small modifications of the Glucose 3.0 SAT solver. The limit for being considered a ``hack''
was arbitrarily set to 1000 non-space character edit distance from the sources of Glucose provided by the organizers. 
Unfortunately, in 2020 there were not enough participants in this sub-track and we do not report on it 
in the results section.

@@

In the Incremental Library track we mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. ``Incremental'' means that
the individual calls to the SAT solver are not independent, but may share 
a common subset of the input clauses or differ in the presence of additional 
unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.

@@

The Parallel track evaluates solver performance in wall-clock time on multiple cores.
The benchmarks and the timeout (5000 seconds) are the same as in the Main track. 
Each solver in this track was run on a m4.16xlarge machine from Amazon Web Services (AWS).
These machines have 64 virtual cores and 256GB of memory. In contrast to the 
Main track, proof logging for unsatisfiable instances is not required in the Parallel track.

@@

New this year was the Cloud track, dedicated to evaluating distributed solvers. Similar to the
Parallel track, the solvers ran on the Amazon cloud, but using 100 machines of the
type m4.4xlarge, which have 16 virtual cores and 64GB of memory. 
Communication between the machines is possible using MPI and SSH.
The wall-clock timeout is 1000 seconds and the same benchmarks were used as in
the Main and Parallel track.

@@

The following requirements were imposed for 
the participation in the competition.

@@

Proofs of unsatisfiability were to be emitted in the DRAT format~\cite{DRATtrim},
either in its textual version, which is also very similar to the DIMACS input format,
or in a more compact binary version (for more details, see \cite{satComp2020www}, Unsat Certificates).
%
Details on certification are further discussed below in Section~\ref{sec:certif}.

\paragraph{Number of submissions}

Each participant was restricted to be an author of at most four different sequential solvers,
two different parallel solvers, and one ``Glucose hack'' sub-track solver.
Two solvers were considered different as soon as their sources differed
or the compilation options were different, or different command line options were used
(with the exception of an option enabling or disabling the proof output).

@@

In all tracks it was required to output a model to certify recognizing a satisfiable instance.
On the other hand, certificates for unsatisfiable instances (proofs) were required only 
in the Main track (besides the No Limits track).
In some cases, a solver output the correct result, but the respective certificate was wrong. 
Such solvers were demoted to the No-limits track of the competition.

The proofs were validated in a two step fashion. First, the tool {\tt DRAT-trim}~\cite{DRATtrim}
was used for initial checking and optimizing the emitted proof, deriving a so-called LRAT proof file.
Afterwards, an independent, formally-verified checker {\tt cake\_lpr} \cite{cakeLprGithub}, 
was used for validating the LRAT proof as a correct proof of unsatisfiability of the given formula.
In the Main track, we only considered solved those unsatisfiable formulas that 
could be validated by {\tt DRAT-trim}.
There were several cases where {\tt cake\_lpr} ran out of resources before being able to 
confirm after {\tt DRAT-trim}. However, there was no case where {\tt DRAT-trim} would accept
a proof and {\tt cake\_lpr} would later disagree with the verdict.

\subsection{Computing environments}

\label{sec:computing}
The Main, No-Limits, and Planning tracks were run on the StarExec cluster \cite{starexec},
whose nodes are equipped with Intel Xeon \SI{2.4}{\giga\hertz} processors 
and \SI{128}{\giga\byte} of memory.
The time limit enforced on each solver for solving an instance was \SI{5000}{\second}. 
(In the Main track, validation then continued for up to \SI{45000}{\second}.)
The solvers were allowed to use the full \SI{128}{\giga\byte} of RAM.\footnote{
Unfortunately, the memory limit of \SI{24}{\giga\byte}, that was used in the previous years,
was by mistake advertised on the competition web page prior to solver submission.
This could have resulted in some solvers not ``daring'' to use the full \SI{128}{\giga\byte}
in the competition.}

@@

The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient. The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch all pairs of adjacent solvers in
rank still gives a value of $\rho = 0.9974$. On the other hand, switching the
highest ranked solver (\solver{Kissat-sat}) with the lowest,
results in a rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to
the bottom of the ranking while moving every other solver up one rank gives a
higher $\rho = 0.8776$. Doing the same to all three \solver{Kissat}
configurations (with ranks 1, 2 and 11, now 46, 47 and 48) results in
$\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances over 50 samples for each size}
   \label{fig:sampleAll}
\end{figure}
The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, still results in an average rank correlation above $0.96$.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\textwidth]{stability/fam_leave_one_out_corr.png}
  \caption{Change of rank correlation under removal of individual benchmark families}
  \label{fig:sampleFamily}
\end{figure}

Figure~\ref{fig:sampleFamily} shows the rank
correlation coefficient resulting from removing a complete benchmark family
instead of a random subset. As expected, removing a non-random subset can have a
higher impact on the ranking even if it is small. Removing the $13$ benchmarks
from the \emph{hgen}-family results in a (still high) rank correlation of
$0.9895$. Additionally, the ranking of the top five solvers stays the same.
The individual removal of all other benchmarks families results in a rank correlation above $0.99$.

>>>>

The same rules for participation as in Main track applied to the Planning track, in which we evaluated the solvers on 200 benchmarks 
which all came from the same application domain -- automated planning. 
Solvers submitted to the Main track automatically participated in the Planning track.
It is envisioned that also in the future there will be an ``application promotion'' track,
each time highlighting a different area where the SAT solving technology helps to advance the state of the art.

Finally, the Main Track also included the so-called ``Glucose hack'' sub-track. 
As in the past several advances in SAT solving required only a small modification of an established solver
to achieve a considerable contribution, this sub-track encouraged participation 
of small modifications of the Glucose 3.0 SAT solver. The limit for being considered a ``hack''
was---somewhat arbitrarily---set to 1000 non-space character edit distance from the sources of Glucose provided by the organizers. 
Unfortunately, in 2020 there were not enough participants in this sub-track and we do not report on it 
in the results section.

@@

In the Incremental Library track we mimic scenarios
where a SAT solver is used as a back-end solver in a more complex tool
(typically solving a harder problem than SAT) and is called multiple times before 
the enclosing tool reaches its final state. \textcolor{red}{Could mention concrete applications here?}
``Incremental'' here refers to the 
idea that
the individual calls to the SAT solver are not independent, but may share 
a common subset of the input clauses or differ in the presence of additional 
unit clause assumptions~\cite{Nadel:2014:Incremental,Fazekas:2019:IncrementalInprocessing}.

@@

The Parallel track evaluates the runtime performance in wall-clock time of SAT solvers making use of
 multiple processor cores.
The benchmarks and the timeout (5000 seconds) are the same as in the Main track \textcolor{red}{repetitive? also mentioned below?}. 
Each solver in this track was run on a m4.16xlarge machine from Amazon Web Services (AWS).
These machines have 64 virtual cores and 256GB of memory. In contrast to the 
Main track, proof logging for unsatisfiable instances is not required in the Parallel track.

@@

New this year was the Cloud track, dedicated to evaluating distributed solvers \textcolor{red}{Not sure
if terms ``distributed'' vs ``parallel'' are clear, could be more explicitly explained?}. Similar to the
Parallel track, the solvers ran on the Amazon cloud, but using 100 machines of the
type m4.4xlarge, which have 16 virtual cores and 64GB of memory. 
Communication between the machines is possible using MPI and SSH.
The wall-clock timeout is 1000 seconds and the same benchmarks were used as in
the Main and Parallel track. \textcolor{red}{repetitive? also mentioned below?}

@@

The following requirements were imposed for 
participating in the competition.

@@

Proofs of unsatisfiability were to be emitted in the DRAT format~\cite{DRATtrim},
either in its textual version, which is also very similar to the DIMACS input format,
or in a more compact binary version (for more details, see \cite{satComp2020www}, Unsat Certificates).
%
Details on certification are further discussed in Section~\ref{sec:certif}.

\paragraph{Number of Submissions}

Each participant was allowed to be an author of at most four different sequential solvers,
two different parallel solvers, and one ``Glucose hack'' sub-track solver.
Two solvers were considered different as soon as their sources differed
or the compilation options were different, or different command line options were used
(with the exception of an option enabling or disabling the proof output).

@@

In all tracks solvers were required to output a model to certify recognizing a satisfiable instance.
On the other hand, 
certificates for unsatisfiable instances (proofs) were required only 
in the Main track (besides the No Limits track).
In some cases, a solver output the correct result, but the respective certificate was wrong. 
Such solvers were demoted to the No-limits track of the competition.

The proofs were validated in a two-step fashion. First, the tool {\tt DRAT-trim}~\cite{DRATtrim}
was used for initial checking and optimizing the emitted proof, deriving a so-called LRAT proof file.
Afterwards, an independent, formally-verified checker {\tt cake\_lpr} \cite{cakeLprGithub}, 
was used for validating the LRAT proof as a correct proof of unsatisfiability of the given formula.
In the Main track, we only considered solved those unsatisfiable formulas that 
could be validated by {\tt DRAT-trim}.
There were several cases where {\tt cake\_lpr} ran out of resources before being able to 
confirm after {\tt DRAT-trim}. However, there was no case where {\tt DRAT-trim} would accept
a proof and {\tt cake\_lpr} would later disagree with the verdict.

\subsection{Computing Environments}

\label{sec:computing}
The Main, No-Limits, and Planning tracks were run on the StarExec cluster \cite{starexec}
with computing nodes  equipped with Intel Xeon \SI{2.4}{\giga\hertz} processors 
and \SI{128}{\giga\byte} of memory.
The time limit enforced on each solver for solving an instance was \SI{5000}{\second}. 
(In the Main track, validation then continued for up to \SI{45000}{\second}.)
The solvers were allowed to use the full \SI{128}{\giga\byte} of RAM.\footnote{
Unfortunately, the memory limit of \SI{24}{\giga\byte}, that was used in the previous years,
was by mistake advertised on the competition web page prior to solver submission.
This could have resulted in some solvers not ``daring'' to use the full \SI{128}{\giga\byte}
in the competition.}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{stability/ALL_random_smpling_correlations.png}
  \caption{Mean and standard deviation of rank correlation under removal of random instances over 50 samples for each size}
   \label{fig:sampleAll}
\end{figure}
The coefficient $\rho$ is in the interval $[-1, 1]$, where a rank correlation of
$1$ means that the two rankings are equal and $-1$ means that one ranking is the
reverse of the other. To give a better intuition for $\rho$, we list a few
modifications to the original ranking together with the resulting rank
correlation coefficient. The smallest change we can make is to switch the rank
of two adjacent solvers, resulting in a high rank correlation $\rho = 0.9999$.
Several small changes also result in a high rank correlation; repeating the same
modification as before $n/2$ times to switch all pairs of adjacent solvers in
rank still gives a value of $\rho = 0.9974$. On the other hand, switching the
highest ranked solver (\solver{Kissat-sat}) with the lowest,
results in a rank correlation of $\rho = 0.7602$. Moving \solver{Kissat-sat} to
the bottom of the ranking while moving every other solver up one rank gives a
higher $\rho = 0.8776$. Doing the same to all three \solver{Kissat}
configurations (with ranks 1, 2 and 11, now 46, 47 and 48) results in
$\rho = 0.6839$.

The mean and standard deviation of the computed correlation coefficients are
depicted in Figure~\ref{fig:sampleAll}.
The rank correlation is high even for relatively small samples.
The average rank correlation drops below $0.99$ only after randomly removing at least $95$ benchmarks, which is $30\%$ of the considered benchmark set. Accordingly, removing fewer benchmarks randomly has almost no effect on the ranking of the solvers.
Furthermore, removing fewer than $200$ ($63\%$) benchmarks, still results in an average rank correlation above $0.96$.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.9\textwidth]{stability/fam_leave_one_out_corr.png}
  \caption{Change of rank correlation under removal of individual benchmark families}
  \label{fig:sampleFamily}
\end{figure}

Figure~\ref{fig:sampleFamily} shows the rank
correlation coefficient resulting from removing a complete benchmark family. As expected, removing a non-random subset can have a
higher impact on the ranking even if it is small. Removing the $13$ \emph{hgen} benchmarks
results in a (still high) rank correlation of
$0.9895$. Additionally, the ranking of the top five solvers stays the same.
The individual removal of all other benchmarks families results in a rank correlation above $0.99$.
